---
title: "Lab 13 - Classification Trees (Solution)"
author: Karthik R Patil
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the packages 
```{r}
library(rpart)
library(rpart.plot)
library(caret)

```

## 1 create a data frame
```{r}
# load the data and set stringsAsFactors to TRUE
ebay.df<-read.csv("eBayAuctions.csv",stringsAsFactors  =  TRUE)
# first six rows 
head(ebay.df)
# column names
names(ebay.df)
```

## 2 convert numeric variables to categorical variables
```{r}
# convert Duration to a categorical variable
ebay.df$Duration <- as.factor(ebay.df$Duration)
# convert Competitive. to a categorical variable
ebay.df$Competitive. <- as.factor(ebay.df$Competitive.)
# return the structure 
str(ebay.df)
```

## 3 partition the data into train (60%) and validation (40%) sets
```{r}
# set the seed 
 # set the seed 
set.seed(1)
# row numbers of the training set 
train.index <- sample(c(1:dim(ebay.df)[1]), dim(ebay.df)[1]*0.6)  
head(train.index)
# training set
# training set
train.df <- ebay.df[train.index, ]
head(train.df)
# test set 
test.df <- ebay.df[-train.index, ]
head(test.df)
```

## 4 fit a classification tree on the training set 
```{r}
# set seed 
set.seed(1)
# classification tree
ct <- rpart(Competitive. ~ ., data = train.df, method = "class",cp=0.00001,minbucket=30)
```

## 5 generate predicted classes for records in the test set 
```{r}
# predicted classes
pred.class<-predict(ct,test.df,type = "class")
# first six values 
head(pred.class)
```

## 6 create a confusion matrix
```{r}
confusionMatrix(pred.class, test.df$Competitive., positive = "1")

```

## 7 prune the tree
```{r}
# display the cp table
ct$cptable
# xerror values 
ct$cptable[,"xerror"]
# index of the row with the minimum xerror
which.min(ct$cptable[,"xerror"])
# CP value of the row with the minimum xerror 
ct$cptable[which.min(ct$cptable[,"xerror"]),"CP"]
```

## 8 identify the best-pruned tree
```{r}
pruned.ct <- prune(ct, cp = ct$cptable[which.min(ct$cptable[,"xerror"]),"CP"])

```

## 9 plot the pruned tree
```{r}
prp(pruned.ct, type = 1, extra = 1)

```

## 10 generate predicted classes for records in the test set 
```{r}
# predicted classes 
pruned.pred.class <- predict(pruned.ct,test.df,type = "class")
# first six values 
head(pruned.pred.class)
```

## 11 create a confusion matrix
```{r}

confusionMatrix(pruned.pred.class, test.df$Competitive., positive = "1")
```