---
title: "Lab-05: Data Visualization"
author: KARTHIK R PATIL
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the package 
```{r}
library(tidyverse)


```

## 1 create a data frame 
```{r}
# read data from a csv file 
housing.df= read.csv("BostonHousing.csv")

# first six rows 
head(housing.df)
# column names 
names(housing.df)
```

## 2 scatter plot

### 2.1 scatter plot of MEDV on LSTAT 
```{r}
ggplot(data = housing.df) + 
geom_point(mapping = aes(x = LSTAT, y = MEDV))

```

### 2.2 scatter plot of MEDV on LSTAT by CAT..MEDV
```{r}

ggplot(data = housing.df)+
  geom_point(mapping = aes(x = LSTAT, y = MEDV, color = as.factor(CAT..MEDV)))

```

## 3 trend line

### 3.1 add a smooth curve that follows the trajectory of the relationship between MEDV and LSTAT 
```{r}
ggplot(data = housing.df, mapping = aes(x = LSTAT, y = MEDV)) + 
  geom_point(color = "navy", alpha = 2/10) +
  geom_smooth(color = "orange", size = 1)


```

### 3.2 add a linear trend line
```{r}
ggplot(data = housing.df, mapping = aes(x = LSTAT, y = MEDV)) + 
  geom_point(color = "navy", alpha = 2/10) +
  geom_smooth(method = "lm", color = "orange", size = 1, se = FALSE)
 
```

## 4 line plot 

### 4.1 fit a simple linear regression model 
```{r}
# use the linear model function to directly compute the trend line that best fits the data
lm <- lm(MEDV ~ LSTAT, data = housing.df )
# summary of regression results  
summary(lm)
# head of the predicted values
head(lm)

```


### 4.2 connect predictions 
```{r}
ggplot(data = housing.df) +
  geom_point(mapping = aes(x = LSTAT, y = MEDV), color = "navy", alpha = 2/10) +
  geom_line(mapping = aes(x = LSTAT, y = lm$fitted.values), color = "orange", size = 1)

```

## 5 sumamry statistics 

### 5.1 visualize average MEDV by CHAS 
```{r}
ggplot(data = housing.df) + 
  stat_summary(mapping = aes(x = CHAS, y = MEDV), geom = "bar", fun = "mean", fill = "blue")

```

### 5.2 visualize average CAT..MEDV by CHAS 
```{r}
ggplot(data = housing.df) + 
  stat_summary(mapping = aes(x = CHAS, y = CAT..MEDV), geom = "bar", fun = "mean", fill = "blue")
```

## 6 histogram

### 6.1 visualize the distribution of MEDV  
```{r}
ggplot(data = housing.df) + 
  geom_histogram(mapping = aes(x = MEDV), binwidth = 5, fill = "blue", color = "white")

```