---
title: "Lab 03 - Data Wrangling"
author: Karthik R Patil
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the package 
```{r}
library(tidyverse)
```

## 1 create a tibble 

```{r}
# load the data 
df<-read_csv("firm.csv")
df
# first six rows 
head(df,6)
# number of rows  
nrow(df)
# number of columns 
ncol(df)
# column names 
names(df)

```

## 2 subset data 

### 2.1 one column 
```{r}
df$NI
select(df,NI)

```

### 2.2 three columns 
```{r}
df[,c("FYEAR","TIC","NI")]
select(df,FYEAR,TIC,NI)

```

### 2.3 four rows and three columns
```{r}
df[c(1:4),c("FYEAR","TIC","NI")]

```

## 3 filter rows  

### 3.1 observations that belong to AAPL in 2015 
```{r}
df[df$FYEAR==2015 & df$TIC== "AAPL",]
filter(df, FYEAR==2015, TIC== "AAPL")
```

### 3.2 observations that have the minimum NI value 
```{r}
df[df$NI==min(df$NI),] 
filter(df,NI==min(NI))
```

### 3.3 select columns of observations that belong to AAPL in 2015   
```{r}
df[df$FYEAR==2015 & df$TIC=="AAPL",c("FYEAR","TIC", "NI")]

```

## 4 arrange rows 

### 4.1 ascending order of NI
```{r}
arrange(df, NI)

```

### 4.2 descending order of NI
```{r}
arrange(df, desc(NI))

```

### 4.3 ascending order of TIC and FYEAR, and descending order of NI
```{r}
arrange(df, TIC, FYEAR, desc(NI))

```

## 5 create new variables  

### 5.1 return on assets 
```{r}
df$ROI <- df$NI/df$AT

```

### 5.2 net income bin numbers  
```{r}
# step1: identify thresholds of 4 bins with an equal number of observations 
bins<- c(quantile(df$NI,0.0), quantile(df$NI, 0.25), quantile(df$NI,0.5),quantile(df$NI,0.75),quantile(df$NI,1) )
bins
# step2: remove the names of a named vector  
bins<-unname(bins)
# step3: bin NI  
df$NI_bin<- .bincode(df$NI, bins,include.lowest= TRUE)

table(df$NI_bin)
```

### 5.3 log of assets 
```{r}
mutate(df, AT_LOG=log(df$AT, base= exp(1)))
head(df,6)

```

### 5.4 drop a column 
```{r}

df$ROI <- NULL
head(df,6)
```

