---
title: "Lab 04 - Data Exploration"
author: Karthik R Patil
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the package 
```{r}
library(tidyverse)
```

## 1 create a tibble 

```{r}
# load the data 
df<- read_csv("firm.csv")

# first six rows 
head(df)

# column names
names(df)
```

## 2 calculate summary statistics 

```{r}
# 2.1 average net income   
mean(df$NI)
# 2.2 first decile of net income  
quantile(df$NI, 0.10)
# 2.3 summary of net income    
summary(df$NI)

```

## 3 compute counts and proportion of logical values    
```{r}
# 3.1 number of rows 
sum(df$NI>10000 & df$DV >0)

# 3.2 proportion of rows 
mean(df$NI>10000 & df$DV >0)
```

## 4 summarize data with dpylr  

### 4.1 average and maximum net income 
```{r}
summarize(df, NI_mean= mean(NI), NI_max= max(NI),n_rows=n(), n_firms=n_distinct(TIC))

```

### 4.2 average and minimum net income by firm 
```{r}

df_by_TIC<- group_by(df, TIC)
summarize(df_by_TIC, NI_mean= mean(NI),NI_min= min(NI), n_rows= n())
```

## 5 combine multiple operations with the pipe 

### 5.1 average net income by firm  
```{r}
df %>%
  group_by(TIC) %>%
  summarize( NI_mean= mean(NI)) %>%
  arrange(desc(NI_mean)) 
  
  
```

### 5.2 most recent return on assets by firm 
```{r}

df$ROA=df$NI/df$AT
df %>% 
  group_by(TIC) %>% 
  arrange(desc(FYEAR)) %>% 
  summarize(recent_ROA=first(ROA))
```

## 6 combine firm fundamentals with executives’ compensation	

### 6.1 load executive compensation data 
```{r}
# load the data 
manager <- read_csv("manager.csv")

# first six rows 
head(manager)
# column names
names(manager)
```

### 6.2  frequency counts of firm
```{r}
table(df$TIC)

```

### 6.3 frequency counts of year
```{r}
table(manager$TICKER)

```

### 6.4 inner join 
```{r}
# inner join 
merged<- inner_join(manager, df, by=c("TICKER"= "TIC", "YEAR"= "FYEAR"))

# first six rows 
head(merged)
# column names 

```

### 6.5 average salary and return on assets during 2017 and 2018 
```{r}

merged %>%
  filter(YEAR==2017 | YEAR==2018) %>%
  group_by (TICKER) %>%
  summarise(SALARY_Mean= mean(SALARY), ROA_mean= mean(ROA),n_rows= n(), .groups= 'drop') %>%
  arrange(desc(ROA_mean))
  
  
```


