---
title: "Lab 08 - Logistic Regression"
author: Karthik R Patil
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the package and suppress scientific notation 
```{r}
# load packages 
library(ggplot2)

# first six rows
ebay.df= 
# not display scientific notation
options(scipen=999)
```

## 1 create a data frame
```{r}
# load the data
ebay.df <- read.csv("eBayAuctions.csv", stringsAsFactors = FALSE)
head(ebay.df)
# variable names 
names(ebay.df)
```

## 2 apply a function to each group 
```{r}
# mean of the binary outcome for each day of the week of auction close   
tapply(ebay.df$Competitive, ebay.df$endDay, mean)

# mean of the binary outcome for each auction category  
tapply(ebay.df$Competitive, ebay.df$Category, mean)
```

## 3 reduce levels in the day of week that the auction closed  
```{r}
# combine Sunday and Friday into a single category called Sun_Fri
ebay.df$endDay== 'Sun'
ebay.df$endDay[ebay.df$endDay=='Sun']<- 'Sun_Fri'
ebay.df$endDay[ebay.df$endDay=='Fri']<- 'Sun_Fri'
```

## 4 reduce levels in auction categories 
```{r}
# combine Business/Industrial and Computers into a single category called Computers
ebay.df$Category[ebay.df$Category=='Business/Industrial']<- 'Computer'

# combine Antique/Art/Craft and Collectibles into a single category called collectibles
ebay.df$Category[ebay.df$Category=='Antique/Art/Craft']<- 'Collectibles'
```

## 5 convert Duration to a categorical or factor variable 
```{r}
# convert Duration from a numeric variable to a categorical or factor variable  
ebay.df$Duration  <- as.factor(ebay.df$Duration)
# variable types 
str(ebay.df$Duration)
```

## 6 data partition
```{r}
# set seed for reproducing the partition 
set.seed(1)
# takes a sample of row numbers for the training set 
train.index <- sample(c(1:dim(ebay.df)[1]), dim(ebay.df)[1]*0.6)  
head(train.index)
# training set 
train.df <- ebay.df[train.index, ]
head(train.df)


# test set 
test.df <- ebay.df[-train.index, ]
head(test.df)

```

## 7 fit a logistic regression model 
```{r,warning=FALSE}
# logistic regression using all the variables
reg <- glm(Competitive.~., data = train.df, family = "binomial") 

# summary table 
summary(reg)
```

## 8 generate predicted probabilities for records in the test set 
```{r}
# predicted probabilities 
pred <- predict(reg, test.df, type = "response")

# first six values 
head(pred)
```

## 9 confusion matrix 
```{r}
library(caret)

confusionMatrix(as.factor(ifelse(pred > 0.5, 1, 0)), 
                as.factor(test.df$Competitive.),positive = "1")



```

## 10 creating a gain table 
```{r}
# gain table 
library(gains)

gain <- gains(test.df$Competitive., pred, groups=10)
# cumulative percentage of competitive auctions 
gain$cume.pct.of.total
# cumulative number of auctions 
gain$cume.obs 
```

## 11 plot a lift chart
```{r}
# plot the cumulative number of competitive auctions against the cumulative number of auctions
plot(c(0,gain$cume.pct.of.total*sum(test.df$Competitive.))~c(0,gain$cume.obs), 
     xlab="cumulative number of actions", ylab="cumulative number of competitive actions", type="l")
# add a baseline curve 
lines(c(0,sum(test.df$Competitive.))~c(0, dim(test.df)[1]))
```